### 2018-01-23 ChIP-seq_ER/qrime/04.QRIMESceneInvestigatonOverview.Rmd

---
title: "QRIME Scene Investigation - Overview"
author: "Andrew Holding & Matt Eldridge"
date: "9 January 2017"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
header-includes:
- \usepackage{longtable}
- \usepackage{booktabs}
---

****************************

```{r configuration, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(cache = FALSE)
```

```{r load_libraries, include = FALSE, cache = FALSE}
library(tidyr)
library(dplyr)
library(tibble)
library(MSnbase)
library(RColorBrewer)
library(ggplot2)
library(kfigr)
library(gridExtra)
library(xtable)
library(limma)
```

```{r load_functions, include = FALSE, cache = FALSE}
source("Functions/plots.R")
source("Functions/normalization.R")
```

```{r xtable_options, include=FALSE, cache = FALSE}

options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)

addtorow          <- list()
addtorow$pos      <- list()
addtorow$pos[[1]] <- c(0)
addtorow$command  <- c(paste("\\hline \n",
                             "\\endhead \n",
                             "\\hline \n",
                             "{\\footnotesize Continued on next page} \n",
                             "\\endfoot \n",
                             "\\endlastfoot \n",sep=""))
```

```{r read_protein_information, include = FALSE}
proteinInfo <- read.delim("ReferenceData/protein_gene_description.txt", stringsAsFactors = FALSE)
```

```{r read_metadata, include = FALSE}
metadata <- read.delim("SampleData/metadata.txt", check.names = FALSE)
contrastsTable <- read.delim("SampleData/contrasts.txt", check.names = FALSE)
runs <- levels(metadata$Run)
groups <- levels(metadata$Group)
controlGroup <- "IgG"
```

```{r colours, include = FALSE}
groupColours <- c(brewer.pal(9, "Blues")[5:7], brewer.pal(9, "RdPu")[5:7], brewer.pal(9, "Greens")[6])
names(groupColours) <- groups
sampleColours <- groupColours[metadata$Group]
names(sampleColours) <- metadata$Sample
```

```{r read_intensities, include = FALSE}

peptideData <- list()

for (run in runs)
{
  cat(run, "\n", sep = "")

  data <- read.delim(paste("ProcessedData/", run, ".txt", sep = ""), stringsAsFactors = FALSE, check.names = FALSE)

  peptideCount <- data %>% select(Sequence, Modifications) %>% distinct %>% nrow
  cat("Peptides:", peptideCount, "\n", sep = "")
  if (peptideCount != nrow(data))
    stop("Error: multiple rows for at least one peptide/modification (is this PSM-level data?)")

  proteinCount <- data %>% select(Protein) %>% distinct %>% nrow
  cat("Proteins:", proteinCount, "\n", sep = "")

  sampleColumns <- metadata %>%
    filter(Run == run) %>%
    select(Group) %>%
    unlist(use.names = FALSE) %>%
    as.character

  data <- MSnSet(
    exprs = data %>% select(one_of(sampleColumns)) %>% as.matrix,
    fData = data %>% select(Sequence, Modifications, Protein),
    pData = metadata %>% filter(Run == run) %>% mutate(RowName = Group) %>% column_to_rownames("RowName")
  )

  peptideData[[run]] <- data
}
```

\newpage

# SampleSheet

SampleSheet and ContrastTable configurations are loaded from tab-delimited format text files in the 'SampleData' folder. Reports are generated where requested and output to the 'Reports' folder. Plain text results are also provided in the 'Results' folder.

```{r samples_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
print(
  xtable(
    metadata %>% select(Run, Tag, Group) %>% spread(Run, Tag),
    caption = "Isobaric tags used for each sample (group) and run."
  ),
  include.rownames = FALSE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)

print(
  xtable(
    contrastsTable %>% select(Contrast, Report),
    caption = "Configured Constrasts."
  ),
  include.rownames = FALSE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```

# Peptide intensity data

```{r filter_missing_intensities, include=FALSE}
peptideDataExcludingMissingValues <- list()
for (run in runs)
{
  cat(run, "\n", sep = "")
  data <- peptideData[[run]]
  data <- data[which(complete.cases(exprs(data))),]
  peptideDataExcludingMissingValues[[run]] <- data
}
```

```{r combine_peptide_intensities, include = FALSE}

# function for combining data for runs into single MSnSet object
combineRuns <- function(runData, runs)
{
  combinedData <- NULL

  for (run in runs)
  {
    data <- runData[[run]]

    sampleNames(data) <- pData(data) %>%
      transmute(Sample = paste(Run, Group, sep = ":")) %>%
      unlist %>%
      as.character

    featureNames(data) <- fData(data) %>%
      transmute(ID = paste(Sequence, Modifications, sep = "|")) %>%
      unlist %>%
      as.character()

    if (is.null(combinedData))
    {
      combinedData <- data
    } else
    {
      combinedData <- BiocGenerics::combine(combinedData, data)
    }
  }

  return(combinedData)
}

combinedPeptideData <- combineRuns(peptideData, runs)
combinedPeptideDataExcludingMissingValues <- combineRuns(peptideDataExcludingMissingValues, runs)
```

```{r peptide_protein_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

countProteins <- function(data) data %>% fData %>% select(Protein) %>% distinct %>% nrow

peptideData[["All"]] <- combinedPeptideData
peptideDataExcludingMissingValues[["All"]] <- combinedPeptideDataExcludingMissingValues

counts <- bind_rows(
  peptideData %>% sapply(nrow) %>% t %>% as.data.frame %>% mutate(label = "Peptides"),
  peptideDataExcludingMissingValues %>% sapply(nrow) %>% t %>% as.data.frame %>% mutate(label = "Peptides with no missing values"),
  peptideData %>% sapply(countProteins) %>% t %>% as.data.frame %>% mutate(label = "Proteins"),
  peptideDataExcludingMissingValues %>% sapply(countProteins) %>% t %>% as.data.frame %>% mutate(label = "Proteins with no missing values")
)
rownames(counts) <- counts$label
counts <- counts %>% select(-label)

peptideData[["All"]] <- NULL
peptideDataExcludingMissingValues[["All"]] <- NULL

print(
  xtable(
    counts,
    align = "lrrrr",
    caption = "Numbers of peptides and proteins observed in each run."
  ),
  include.rownames = TRUE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```

`r figr("intensity_plots", TRUE, type="Figure")` shows the distribution of intensities for each sample within each run.

```{r intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of intensities from each run.", fig.height=8, fig.width=5}

minIntensity <- -2
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideData[[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```

\newpage

# Normalization of intensity data

```{r imputation, include = FALSE}
peptideDataImputedMin <- list()
peptideDataImputedKNN <- list()
for (run in runs)
{
  peptideDataImputedMin[[run]] <- impute(peptideData[[run]], "min")
  peptideDataImputedKNN[[run]] <- impute(peptideData[[run]], "knn")
}
```

```{r normalization, include = FALSE}

peptideDataNormalized <- list()

peptideDataNormalized[["IncludeMissingValues_NoNormalization"]] <- peptideData
peptideDataNormalized[["ExcludeMissingValues_NoNormalization"]] <- peptideDataExcludingMissingValues
peptideDataNormalized[["ImputeMin_NoNormalization"]] <- peptideDataImputedMin
peptideDataNormalized[["ImputeKNN_NoNormalization"]] <- peptideDataImputedKNN

for (missingValueMethod in c("IncludeMissingValues", "ExcludeMissingValues", "ImputeMin", "ImputeKNN"))
{
  peptideDataForNormalization <- peptideDataNormalized[[paste(missingValueMethod, "NoNormalization", sep = "_")]]

  quantileNormalizationAnalysis <- paste(missingValueMethod, "QuantileNormalization", sep = "_")
  quantileNormalizationExcludingIgGAnalysis <-
    paste(missingValueMethod, "QuantileNormalizationExcludingIgG", sep = "_")
  scaleNormalizationAnalysis <- paste(missingValueMethod, "ScaleNormalization", sep = "_")
  scaleNormalizationExcludingIgGAnalysis <-
    paste(missingValueMethod, "ScaleNormalizationExcludingIgG", sep = "_")
  scaleNormalizationTopIgGAnalysis <- paste(missingValueMethod, "ScaleNormalizationTopIgG", sep = "_")

  peptideDataNormalized[[quantileNormalizationAnalysis]] <- list()
  peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationExcludingIgGAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationTopIgGAnalysis]] <- list()

  numberOfPeptidesWithHighestIgG <- 10

  for (run in runs)
  {
    data <- peptideDataForNormalization[[run]]

    # quantile normalization
    peptideDataNormalized[[quantileNormalizationAnalysis]][[run]] <- normalize(data, method = "quantiles")

    # scale normalization
    peptideDataNormalized[[scaleNormalizationAnalysis]][[run]] <- normalizeMedianScaling(data)

    samples <- setdiff(sampleNames(data), controlGroup)

    # quantile normalization
    normalized <- normalize(data[,samples], method = "quantiles")
    peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]][[run]] <- data
    exprs(peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]][[run]])[,samples] <-
      exprs(normalized)

    # scale normalization
    peptideDataNormalized[[scaleNormalizationExcludingIgGAnalysis]][[run]] <-
      normalizeMedianScaling(data, samples)

    # scale normalization using peptides with highest IgG intensities for computing sample medians
    controlIntensities <- data %>%
      exprs %>%
      as.data.frame %>%
      select(one_of(controlGroup)) %>%
      unlist %>%
      as.numeric
    controlIntensities[is.na(controlIntensities)] <- -Inf
    controlIntensityOrder <- rev(order(controlIntensities))
    peptideDataNormalized[[scaleNormalizationTopIgGAnalysis]][[run]] <-
      normalizeMedianScaling(data, rowsForCalculatingMedian = controlIntensityOrder[1:numberOfPeptidesWithHighestIgG])
  }
}
```

```{r normalized_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where quantile normalization was applied to peptide intensities that include misssing values.", fig.height=6.5, fig.width=5}

analysis <- "IncludeMissingValues_QuantileNormalization"

minIntensity <- 0
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```

\newpage

# Protein-level quantification

```{r summarize_protein_intensities, include=FALSE}

proteinData <- list()

for (analysis in names(peptideDataNormalized))
{
  peptideDataForSummarization <- peptideDataNormalized[[analysis]]

  proteinData[[analysis]] <- list()

  for (run in runs)
  {
    data <- peptideDataForSummarization[[run]]

    # exclude rows with missing values (will have no effect if imputation carried out)
    data <- data[which(complete.cases(exprs(data))),]

    data <- combineFeatures(data, groupBy = fData(data)$Protein, fun = "sum", na.rm = FALSE)
    fData(data) <- fData(data) %>% select(Protein)

    proteinData[[analysis]][[run]] <- data
  }
}
```

```{r combine_protein_intensities, include = FALSE}

combinedProteinData <- list()

for (analysis in names(proteinData))
{
  for (run in runs)
  {
    data <- proteinData[[analysis]][[run]]

    sampleNames(data) <- pData(data) %>%
      transmute(Sample = paste(Run, Group, sep = ":")) %>%
      unlist %>%
      as.character

    if (is.null(combinedProteinData[[analysis]]))
    {
      combinedProteinData[[analysis]] <- data
    } else
    {
      combinedProteinData[[analysis]] <- BiocGenerics::combine(combinedProteinData[[analysis]], data)
    }
  }
}
```

```{r replicate_counts, include=FALSE}

isRepresented <- function(x) any(!is.na(x))

getReplicateCounts <- function(data)
{
  metadata <- pData(data) %>% rownames_to_column("Sample")

  intensities <- data %>%
    exprs %>%
    as.data.frame %>%
    mutate(N = 0)

  for (run in runs)
  {
    samples <- metadata %>%
      filter(Run == run) %>%
      select(Sample) %>%
      unlist %>%
      as.character

    intensities <- intensities %>%
      mutate(N = N + apply(select(., one_of(samples)), 1, isRepresented))
  }

  replicateCounts <- intensities %>%
    select(N) %>%
    mutate(Count = 0) %>%
    group_by(N) %>%
    summarize_each(funs(length)) %>%
    ungroup %>%
    mutate(N = as.character(N))

  replicateCounts <- replicateCounts %>%
    bind_rows(data_frame(N = "total", Count = sum(replicateCounts$Count)))

  return(replicateCounts)
}

replicateCounts <- getReplicateCounts(combinedProteinData[["ImputeKNN_ScaleNormalization"]])

replicateCountTableData <- inner_join(
  getReplicateCounts(combinedProteinData[["ImputeKNN_ScaleNormalization"]]) %>%
    rename(`Including missing values` = Count),
  getReplicateCounts(combinedProteinData[["ExcludeMissingValues_ScaleNormalization"]]) %>%
    rename(`Excluding missing values` = Count),
  by = "N"
)

replicateCountTableData <- replicateCountTableData %>%
  as.data.frame %>%
  column_to_rownames(var = "N") %>%
  t %>%
  as.data.frame %>%
  rownames_to_column(var = "Runs/replicates")
```

```{r replicate_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# colnames(replicateCounts) <- c("Runs/replicates", "Proteins (includes missing values)", "Proteins")
print(
  xtable(
    replicateCountTableData,
    # align = "llr",
    caption = "Numbers of proteins identified in differing numbers of runs and total number of proteins identified in all runs."
  ),
  include.rownames = FALSE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```

# Principal Component Analysis

```{r pca_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for proteins sampled in all 3 runs. The PCA was based on protein-level data resulting from summation of quantile normalized peptide intensities in which missing values were imputed using KNN-based nearest neighbour averaging. The first two principal components are displayed.", fig.height=5}

data <- combinedProteinData[["ImputeKNN_QuantileNormalization"]]

metadata <- pData(data) %>% rownames_to_column("Sample")

scaledIntensities <- data %>%
  exprs %>%
  as.data.frame %>%
  filter(complete.cases(.))

for (run in runs)
{
  samples <- metadata %>%
    filter(Run == run) %>%
    select(Sample) %>%
    unlist %>%
    as.character

  summedIntensities <- scaledIntensities %>%
    select(one_of(samples)) %>%
    rowSums

  scaledIntensities <- scaledIntensities %>%
    mutate_each(funs(. / summedIntensities), one_of(samples))
}

print(pcaPlot(scaledIntensities, metadata$Sample, metadata$Group, groupColours, labels = metadata$Run, legend = TRUE))
```

```{r knitreports, include = FALSE, cache = FALSE}

library("rmarkdown")

for (contrast in contrastsTable$Contrast[contrastsTable$Report=="Y"])
{
  render('05.QRIMESceneInvestigatonReport.Rmd', output_file = paste('Reports/05.QRIMEReport_', contrast, '.pdf', sep = ''))
}

```

```{r save, eval = TRUE, include = FALSE, cache = FALSE}
save.image(file = "Results/QRIME.RData")
```