### 2018-01-23 Manuscript_Figures_Combined.Rmd

---
title: "VULCAN Analysis Combined"
author: "Andrew Holding"
date: "1/22/2018"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: tango
    number_sections: yes
    toc: yes
    toc_depth: 3
  html_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
always_allow_html: yes
---

```{r setup_master, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

#Estrogen Receptor-alpha qPCR Analysis

```{r ER qRPCR, child = 'qPCR_ER/Manuscript_Figures.Rmd'}
```

#Estrogen Receptor-alpha ChIP-Seq and VULCAN Analysis

```{r ER ChIP-seq, child = 'ChIP-seq_ER/rmarkdown.Rmd'}
```

#Estrogen Receptor-alpha qPLEX-RIME Analysis

```{r ER qRIME, child = 'qPLEX-RIME_ER/02.Supplementary.Rmd'}
```

#GRHL2 ChIP-Seq Analysis

```{r GRHL2 Chip, child = 'ChIP-seq_GRHL2/Manuscript_Figures.Rmd'}
```

#GRHL2 qPLEX-RIME Analysis

```{r GRHL2 qRIME, child = 'qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd'}
```

# Technical Session Info
The following code describes the R environment used to generate this document and will help making it fully reproducible should there be future updates in any of the packages.
```{r sessionInfo}
sessionInfo()
```