### 2018-01-23 qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd

---
title: |
  | Proteomics RIME-TMT analysis report
  | Holding: Changes in GRHL2 interactome on Activation
always_allow_html: yes
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: '3'

---

```{r setup_qplex, include=FALSE}

library(UniProt.ws)
library(plyr)
library(dplyr)
library(magrittr)
library(qPLEXanalyzer)
library(knitr)
library(readxl)
library(grid)
library(ggplot2)
library(DT)
library(rmarkdown)
options(stringsAsFactors=F)

# knitr options
opts_knit$set(eval.after = 'fig.cap')
```

```{r numberingFunctions, include=F}
# This deals with captioning and links to figure and tables
# Autonumbering is handled automatically in PDFs, but with HTML we need a function to deal with it

# reflink generates links to figures/tables.
# refcaption creates an anchor and a caption for a table/figure
# the "refName" argument matches these together
# e.g. reflink("myFig") will create a link in the text to the figure with a caption created
#   by refcaption("myFig", "This is the caption")
# add refType="Table" to link/caption a table

numberingList <- list(Figure=list(startFig=0), Table=list(startTab=0))

getNumber <- function(refName, refType, numList=numberingList){
    if(!refName%in%names(numList[[refType]])){
        numList[[refType]][[refName]] <- max(unlist(numList[[refType]]))+1 
        numberingList <<- numList
    }
    refNum <- numList[[refType]][[refName]]
    return(refNum)
}
reflink <- function(refName, refType="Figure"){
    refNum <- getNumber(refName, refType)
    paste0("[", refType, " ", refNum, "](#", refName, ")")
}
refcaption <- function(refName, refCaption, refType="Figure"){
    refNum <- getNumber(refName, refType)
    if(refType=="Table"){
        cap <- htmltools::tags$caption(style = 'caption-side: top; text-align: center;',
                                       htmltools::strong(paste0('Table ', refNum, ': ', refCaption)))
    }else{
        cap <- paste0("**", refType, " ", refNum, ". ", refCaption, "  **")
    }
    cat(paste0("<a id=\"", refName, "\"></a>"))
    return(cap)
}

```

```{r myCovPlotFunct, include=F}
coveragePlot2 <- function(data, ProteinID, ProteinName, fastaFile, myCol="brown")
{
    suppressPackageStartupMessages(library(GenomicRanges))
    suppressPackageStartupMessages(library(Biostrings))

    if(!is(data,"MSnSet"))
    stop('data has to be of class MSnSet..')
    if(!is(ProteinID,"character"))
    stop('ProteinID has to be of class character..')
    if(!is(ProteinName,"character"))
    stop('ProteinName has to be of class character..')
    if(!is(fastaFile,"character"))
    stop('fastaFile has to be of class character..')
    if(!is(myCol,"character"))
    stop('myCol has to be of class character..')
    ProteinName <- as.character(ProteinName)

    # extract peptide sequence from MsnSet object
    peptide_seq <- fData(data) %>% 
        filter(Master.Protein.Accessions==ProteinID) %>% 
        use_series(Annotated.Sequence) %>% 
        gsub("^.{4}([A-Z]+).{4}$", "\\1", .)

    ## read protein sequence from fastafile
    Protein_seq <- readAAStringSet(fastaFile)

    ## match peptide sequence with protein sequence and store co-ordinates
    getPosition <- function(peptideSeq, ProteinSeq=Protein_seq){
        vmatchPattern(peptideSeq, ProteinSeq) %>% 
            as.data.frame() %>% 
            dplyr::select(start, end) %>% 
            return()
    }

    features <- lapply(peptide_seq, getPosition) %>% 
      do.call(rbind, .)

    # get percent coverage 
    protWidth <- width(Protein_seq)
    coverage <- GRanges("feature",IRanges(features$start,features$end)) %>% reduce() %>% width() %>% sum()
    Perct <- round(coverage/protWidth*100,2)
    SubTitle <- paste0("Number of Unique Peptides: ", length(unique(peptide_seq)), "\n% Coverage: ", Perct)

    nTicks <- min(c(7, ceiling(protWidth/50) + 1))
    brkTicks <- round(seq(0,protWidth, length.out = nTicks), 0)

    features %>% 
      distinct() %>% 
      ggplot() +
          geom_rect(aes(xmin=start-1, xmax=end, ymin=0, ymax=10), fill=myCol) +
          geom_rect(xmin=0, xmax=protWidth, ymin=0, ymax=10, colour="black", fill=NA, size=0.15) +
          labs(title=ProteinName,
               subtitle=SubTitle) +
          theme(axis.title.y=element_blank(),
                axis.text.y=element_blank(),
                axis.ticks.y=element_blank(),
                panel.background = element_rect(fill = "white"), 
                panel.border = element_blank(),
                plot.subtitle = element_text(hjust = 0.5),
                plot.title = element_text(hjust = 0.5)) +
          scale_x_continuous(limits=c(0,protWidth), breaks=brkTicks) +
          scale_y_continuous(limits=c(0,10), breaks=c(0,10), expand = c(0, 0)) %>% 
    return()
}

```

<!--
Project Title: Holding: Changes in GRHL2 interactome on Activation
Redmine number: 5045
Proteomics LIMS:
    Project Name: GRHL2 0-45 minutes
    Project ID: PR882
    Quantification method: TMT
Reagent used: 11-plex
-->

```{r getData, include=F}

# constants
baitProtein <- "GRHL2"
UniProtAccession <- "Q6ISB3"
intensitiesFile <- "intensities/Velos_PR882_ANH_07122017_Peptide_Intensities.xlsx"
outputDir <- "results/"

# read the intensities excel file and replace NA in the IgG columns with 0.001
intensities <-  read_xlsx(intensitiesFile, 1) %>% 
    mutate_at(vars(starts_with("IGG")), funs(ifelse(is.na(.), 0.001, .)))

# the sample names used by proteomics start with numbers, so R has adds an X to the beginning of the 
# column names. Change the column names to so the "0min" is "Ctrl" "45min" is "ER"
# also "read_xlsx" leaves spaces in the column names, need to replace these with "."
colnames(intensities) <- gsub(" ", ".", gsub("X*0min", "Ctrl", gsub("X*45min", "ER", colnames(intensities))))

# create a metadata table
metadata <- data.frame(Experiment=colnames(intensities)[7:17],
                       Label=factor(c(rep("Ctrl", 4), rep("ER", 4), rep("IgG", 3))),
                       Bio.Rep=c(rep(1:4, 2), 1:3),
                       Tech.Rep=NA,
                       Analyt.Rep=NA,
                       Grp=c(rep(1, 8), rep(2, 3)))

# create MSnset object
rawMSnset <- convertToMSnset(data = intensities, metadata = metadata, 
                             indExpData = 7:17, indFData = 1:6,
                             rmMissing=TRUE)

# fasta file of protein sequence
fastaFile <- paste0("accessoryFiles/", UniProtAccession, ".fasta")
fastaUrl <- paste0("http://www.uniprot.org/uniprot/", UniProtAccession, ".fasta")
download.file(fastaUrl, fastaFile, "wget")

# annotation table
humanUP <- UniProt.ws::UniProt.ws(taxId=9606)
keys <- unique(intensities$Master.Protein.Accessions)
columns <- c("ENTRY-NAME", "PROTEIN-NAMES", "GENES")
annot <- UniProt.ws::select(humanUP, keys, columns, "UNIPROTKB") %>% 
    mutate(GeneSymbol=gsub(" .*", "", GENES)) %>% 
    dplyr::select(Protein="UNIPROTKB", Gene="ENTRY-NAME", Description="PROTEIN-NAMES", GeneSymbol)
```

```{r dataForIntroduction, include=F}

# count protein peptide representation
numberOfSamples <- nrow(metadata)
numberOfPeptides <- nrow(intensities)
numberOfProteins <- length(unique(intensities$Master.Protein.Accessions))

peptidesPerProtein <- table(intensities$Master.Protein.Accessions)
```

## Introduction

This report provides a summary of the results of your proteomics experiment.

Your experiment has **`r numberOfSamples`** samples. **`r numberOfPeptides`** peptides from **`r numberOfProteins`** 
unique proteins were captured in the experiment. Each protein is represented by **`r min(peptidesPerProtein)`** to **`r max(peptidesPerProtein)`**
peptides, but median number of peptides per protein was **`r median(peptidesPerProtein)`**.

## Raw data QC

### Coverage plot

`r reflink("covFig")` shows the coverage of the bait protein, `r baitProtein`, in terms of peptides detected.

```{r coveragePlot, echo=F, fig.width=10, fig.height=2, fig.cap=covCap, results="asis"}
covCap <- refcaption("covFig", paste0("Plot of peptide coverage for ", baitProtein, " (UniprotID: ", UniProtAccession, ")"))
coveragePlot2(data=rawMSnset, ProteinID=UniProtAccession, ProteinName=baitProtein, fastaFile=fastaFile)
```

### Intensity Plot

`r reflink("intFig")` shows the distribution of raw peptide intensities for each sample.

```{r rawIntensityPlot, echo=F, fig.width=10, fig.cap=intCap, results="asis"}
intCap <- refcaption("intFig", "Raw intensities plot.")
sampleColours <- assignColours(rawMSnset, groupColourSchemes = 
                                 c("Blues","Reds","Greens"))
rawIP <- intensityPlot(rawMSnset, sampleColours, title = "Peptide intensity distribution")
print(rawIP)
```

### Peptide intensities for GHRL2

`r reflink("baitIntFig")` shows the raw intensities for each peptide detected for the bait protein `r baitProtein`
in each sample.

```{r baitPeptideIntensities, echo=F, fig.width=10, fig.cap=baitIntCap, results="asis"}
baitIntCap <- refcaption("baitIntFig", "Peptide intensities for GRHL2.")
peptideIntensityPlot(rawMSnset, combinedIntensities=NULL, UniProtAccession,
                                 title = "", ylab = "log2(peptide intensity)",
                                 minIntensity = NA, maxIntensity = NA,
                                 selectedSequence = NULL, selectedModifications = NULL)
```

### Correlation Plot

`r reflink("corFig")` shows a correlation matrix to visualize the level of linear association of samples
within and between groups based on the raw peptide intensities.

```{r rawCorrPlot, echo=F, fig.width=10, fig.cap=corCap, results="asis"}
corCap <- refcaption("corFig", "Correlation plot based on raw intensities.")
corrPlot(rawMSnset, method="shade")
```

### Hierichical clustering dendrogram

`r reflink("hcFig")` shows a dendrogram displaying the hierarchical relationship among samples. The vertical axis 
shows the dissimilarity (measured by  means  of  the  Euclidean  distance)  between  samples:  similar  
samples  appear  on  the  same  branches. Colors  correspond  to sample groups.

```{r rawHCPlot, echo=F, fig.width=10, fig.cap=hcCap, results="asis"}
hcCap <- refcaption("hcFig", "Hierachical clustering based on raw intensities.")
hcTmp <- rawMSnset
exprs(hcTmp) <- exprs(hcTmp)+0.01
label_color <- c(rep("red",4),rep("darkgreen",4),rep("orange",2))
hierarchicalPlot(hcTmp,label_color,branchlength=95,title="Clustering plot")
rm(hcTmp)

```

### PCA Plot

`r reflink("pcFig")` shows a visual representation of the scaled loading of the first two dimensions of a principle
component analysis of the raw peptide intensities.

```{r rawPCAPlot, echo=F, fig.width=10, fig.cap=pcCap, results="asis"}
pcCap <- refcaption("pcFig", "Principle Component Analysis based on raw intensities.")
grpcolor <- c("red", "darkgreen", "orange")
rawPCP <- qPLEXanalyzer::pcaPlot(rawMSnset, groupColours=grpcolor, title = "PCA plot", 
        labels = rawMSnset$Bio.Rep, labelsize=2, legend=TRUE)# + 
    theme(aspect.ratio=1)
print(rawPCP)
```

### QC Conclusion

The sample Ctrl rep 1 appears to be an outlier. The differential anlaysis was carried out twice, first
with this sample included and then again with the sample excluded.

## Full data set analysis

The following section shows the results for analysis of all 8 samples.

```{r analysisFull, include=F}
# normalisation
normFullMSnset <- groupScaling(rawMSnset,func=median, Grp="Grp")
pnormFullMSnset <- summarizeIntensities(normFullMSnset, sum, annot)

# differential analysis
contrasts <- c(ER_v_Ctrl = "ER - Ctrl")
contrastName <- gsub(" - ", "_v_", contrasts)
diffstats <- computeDiffStats(data=pnormFullMSnset, contrasts=contrasts, 
                              applyLog2Transform = TRUE)
diffexpFull <- getContrastResults(diffstats=diffstats, contrast=contrasts, 
                              controlGroup = "IgG", ann = annot) %T>% 
    write.csv(paste0(outputDir, "/DifferentialAbundance.", contrastName, ".FullData.csv"), row.names=F)
```

### Effect of within group normalisation

Normalisation was carried out using median scaling. The experimental samples and the IgG control samples
were normalised separately.
`r reflink("fullQCFig")` show the effects of normalisation on the intensity plots and the principle component
analysis.

```{r withinGroupNormaliation, echo=F, fig.width=10, fig.cap=fullQCCap, results="asis"}
fullQCCap <- refcaption("fullQCFig", "Effects of normalisation on peptide intensities.")
normIP <- intensityPlot(normFullMSnset, sampleColours, title = "WithinGrp Scaling") + 
    theme(legend.position = "bottom")
rawPCP <- qPLEXanalyzer::pcaPlot(normFullMSnset, groupColours=grpcolor, title = "PCA plot", 
        labels = rawMSnset$Bio.Rep, labelsize=2, legend=TRUE) +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(normIP, vp = define_region(row = 1, col = 1)) 
print(rawPCP, vp = define_region(row = 1, col = 2))

```

### Differential analysis results

`r reflink("fullDATab")` shows differential abundancy results. No proteins were statistically differentially abundant.
`r reflink("fullQCFig")` shows an MA plot and a volcano plot of the differental analysis results.

```{r daResTabFull, echo=F, results="asis"}

deFColnames <- c("Protein", "Gene", "Description", "Gene Symbol", "log2(Fold Change)", "Adjusted p-value")
deFCap <- refcaption("fullDATab", "Differential abundancy results using all 8 samples.", refType = "Table")
deFOptions <- list(columnDefs = list(list(className = 'dt-center', targets="_all")), dom='ltip')
diffexpFull %>% 
    dplyr::select(Protein, Gene, Description, GeneSymbol, logFC, adj.P.Val) %>% 
    datatable(colnames=deFColnames, options=deFOptions, rownames = FALSE, caption=deFCap)

```

```{r resPlotFull, echo=F, fig.width=10, fig.cap=fullDACap, results="asis"}
fullDACap <- refcaption("fullDAFig", "MA plot and volcano plot of differential protein abundancy.")

MAPlot <- qPLEXanalyzer::maPlot(diffexpFull, significanceLevel = 0.0025, minLogFoldChangeForLabelling = 1.5,
             title="")  +
          labs(y="log2(Fold Change)")

volPlot <- diffexpFull %>% 
    mutate(phredAdjPval=-log10(adj.P.Val)) %>% 
    mutate(Signif=ifelse(adj.P.Val<=0.01, "<=0.01", ifelse(adj.P.Val<=0.05, "<=0.05", ">0.05"))) %>% 
    ggplot() +
        geom_point(aes(x=logFC, y=phredAdjPval, fill=Signif), colour="black", shape=21) +
        scale_fill_manual(values=c("red", "orange", "black"), name="Adjusted p-value", limits=c("<=0.01", "<=0.05", ">0.05"), drop = TRUE) +
        theme_minimal() +
        labs(x="log2(Fold Change)", y="-log10(Adjusted p-value)") +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(MAPlot, vp = define_region(row = 1, col = 1)) 
print(volPlot, vp = define_region(row = 1, col = 2))

```

## Remove sample Ctrl 1

The following section shows the same analysis, but with the Ctrl rep 1 sample removed

```{r analysisTrim, include=F}
trimMSnset <- rawMSnset[,-1]
sampleColours <- assignColours(trimMSnset, groupColourSchemes = c("Blues","Reds","Greens"))

# normalisation
normTrimMSnset <- groupScaling(trimMSnset, func=median, Grp="Grp")

# summarise
pnormTrimMSnset <- summarizeIntensities(normTrimMSnset, sum, annot)
```

### Effect of within group normalisation

Normalisation was carried out using median scaling. The experimental samples and the IgG control samples
were normalised separately.
`r reflink("trimQCFig")` show the effects of normalisation on the intensity plots and the principle component
analysis.

```{r withinGroupNormaliationTrim, echo=F, fig.width=10, fig.cap=trimQCCap, results="asis"}
trimQCCap <- refcaption("trimQCFig", "Effects of normalisation on peptide intensities.")
normIP <- intensityPlot(normTrimMSnset, sampleColours, title = "WithinGrp Scaling") +
    theme(legend.position = "bottom")
rawPCP <- qPLEXanalyzer::pcaPlot(normTrimMSnset, groupColours=grpcolor, title = "PCA plot", 
        labels = trimMSnset$Bio.Rep, labelsize=2, legend=TRUE) +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(normIP, vp = define_region(row = 1, col = 1)) 
print(rawPCP, vp = define_region(row = 1, col = 2))

```

### ER - Ctrl

```{r contrastERCtrl, include=F}
# differential analysis
contrasts <- c(ER_v_Ctrl = "ER - Ctrl")
contrastName <- gsub(" - ", "_v_", contrasts)
diffstats <- computeDiffStats(data=pnormTrimMSnset, contrasts=contrasts, 
                              applyLog2Transform = TRUE)
diffexpTrim <- getContrastResults(diffstats=diffstats, contrast=contrasts, 
                              controlGroup = "IgG", ann = annot) %T>% 
    write.csv(paste0(outputDir, "/DifferentialAbundance.", contrastName, ".SevenSamples.csv"), row.names=F)
```

`r reflink("trimDATab")` shows differential abundancy results for the contrasts `r contrastName`.
`r reflink("trimQCFig")` shows an MA plot and a volcano plot of the differental analysis results

```{r daResTabTrim, echo=F, results="asis"}

deFCap <- refcaption("trimDATab", "Differential abundancy results using all 8 samples.", refType = "Table")
deFColnames <- c("Protein", "Gene", "Description", "Gene Symbol", "log(Fold Change)", "Adjusted p-value")
deFOptions <- list(columnDefs = list(list(className = 'dt-center', targets="_all")), dom='ltip')
diffexpTrim %>% 
    dplyr::select(Protein, Gene, Description, GeneSymbol, logFC, adj.P.Val) %>% 
    datatable(colnames=deFColnames, options=deFOptions, rownames = FALSE, caption=deFCap)

```

```{r resPlotTrim, echo=F, fig.width=10, fig.cap=trimDACap, results="asis"}
trimDACap <- refcaption("trimDAFig", "MA plot and volcano plot of differential protein abundancy.")

MAPlot <- qPLEXanalyzer::maPlot(diffexpTrim, significanceLevel = 0.0025, minLogFoldChangeForLabelling = 1.5,
             title=names(contrasts)) +
          labs(y="log2(Fold Change)")

volPlot <- diffexpTrim %>% 
    mutate(phredAdjPval=-log10(adj.P.Val)) %>% 
    mutate(Signif=ifelse(adj.P.Val<=0.01, "<=0.01", ifelse(adj.P.Val<=0.05, "<=0.05", ">0.05"))) %>% 
    ggplot() +
        geom_point(aes(x=logFC, y=phredAdjPval, fill=Signif), colour="black", shape=21) +
        scale_fill_manual(values=c("red", "orange", "black"), name="Adjusted p-value", limits=c("<=0.01", "<=0.05", ">0.05")) +
        theme_minimal() +
        labs(x="log2(Fold Change)", y="-log10(Adjusted p-value)") +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(MAPlot, vp = define_region(row = 1, col = 1)) 
print(volPlot, vp = define_region(row = 1, col = 2))

```

### Ctrl - IgG

```{r contrastCtrlIgG, include=F}
# differential analysis
contrasts <- c(Ctrl_v_IgG = "Ctrl - IgG")
contrastName <- gsub(" - ", "_v_", contrasts)
diffstats <- computeDiffStats(data=pnormTrimMSnset, contrasts=contrasts, 
                              applyLog2Transform = TRUE)
diffexpCtrlIgG <- getContrastResults(diffstats=diffstats, contrast=contrasts, 
                              controlGroup = NULL, ann = annot) %T>% 
    write.csv(paste0(outputDir, "/DifferentialAbundance.", contrastName, ".SevenSamples.csv"), row.names=F)
```

`r reflink("CtrlIgGDATab")` shows differential abundancy results for the contrasts `r contrastName`.
`r reflink("CtrlIgGQCFig")` shows an MA plot and a volcano plot of the differental analysis results.

```{r daResTabCtrlIgG, echo=F, results="asis"}

deFCap <- refcaption("CtrlIgGDATab", "Differential abundancy results using all 8 samples.", refType = "Table")
deFColnames <- c("Protein", "Gene", "Description", "Gene Symbol", "log(Fold Change)", "Adjusted p-value")
deFOptions <- list(columnDefs = list(list(className = 'dt-center', targets="_all")), dom='ltip')
diffexpCtrlIgG %>% 
    dplyr::select(Protein, Gene, Description, GeneSymbol, logFC, adj.P.Val) %>% 
    datatable(colnames=deFColnames, options=deFOptions, rownames = FALSE, caption=deFCap)

```

```{r resPlotCtrlIgG, echo=F, fig.width=10, fig.cap=CtrlIgGDACap, results="asis"}
CtrlIgGDACap <- refcaption("CtrlIgGDAFig", "MA plot and volcano plot of differential protein abundancy.")

MAPlot <- qPLEXanalyzer::maPlot(diffexpCtrlIgG, significanceLevel = 0.00, minLogFoldChangeForLabelling = 1000,
                 title=names(contrasts), ylab="log2(Fold Change)")

volPlot <- diffexpCtrlIgG %>% 
    mutate(phredAdjPval=-log10(adj.P.Val)) %>% 
    mutate(Signif=ifelse(adj.P.Val<=0.01, "<=0.01", ifelse(adj.P.Val<=0.05, "<=0.05", ">0.05"))) %>% 
    ggplot() +
        geom_point(aes(x=logFC, y=phredAdjPval, fill=Signif), colour="black", shape=21) +
        scale_fill_manual(values=c("red", "orange", "black"), name="Adjusted p-value", limits=c("<=0.01", "<=0.05", ">0.05")) +
        theme_minimal() +
        labs(x="log2(Fold Change)", y="-log10(Adjusted p-value)") +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(MAPlot, vp = define_region(row = 1, col = 1)) 
print(volPlot, vp = define_region(row = 1, col = 2))

```

### ER - IgG

```{r contrastERIgG, include=F}
# differential analysis
contrasts <- c(ER_v_IgG = "ER - IgG")
contrastName <- gsub(" - ", "_v_", contrasts)
diffstats <- computeDiffStats(data=pnormTrimMSnset, contrasts=contrasts, 
                              applyLog2Transform = TRUE)
diffexpERIgG <- getContrastResults(diffstats=diffstats, contrast=contrasts, 
                              controlGroup = NULL, ann = annot) %T>% 
    write.csv(paste0(outputDir, "/DifferentialAbundance.", contrastName, ".SevenSamples.csv"), row.names=F)
```

`r reflink("ERIgGDATab")` shows differential abundancy results for the contrasts `r contrastName`.
`r reflink("ERIgGQCFig")` shows an MA plot and a volcano plot of the differental analysis results.

```{r daResTabERIgG, echo=F, results="asis"}

deFCap <- refcaption("ERIgGDATab", "Differential abundancy results using all 8 samples.", refType = "Table")
deFColnames <- c("Protein", "Gene", "Description", "Gene Symbol", "log(Fold Change)", "Adjusted p-value")
deFOptions <- list(columnDefs = list(list(className = 'dt-center', targets="_all")), dom='ltip')
diffexpERIgG %>% 
    dplyr::select(Protein, Gene, Description, GeneSymbol, logFC, adj.P.Val) %>% 
    datatable(colnames=deFColnames, options=deFOptions, rownames = FALSE, caption=deFCap)

```

```{r resPlotERIgG, echo=F, fig.width=10, fig.cap=ERIgGDACap, results="asis"}
ERIgGDACap <- refcaption("ERIgGDAFig", "MA plot and volcano plot of differential protein abundancy.")

MAPlot <- qPLEXanalyzer::maPlot(diffexpERIgG, significanceLevel = 0.00, minLogFoldChangeForLabelling = 1000,
                 title=names(contrasts), ylab="log2(Fold Change)")

volPlot <- diffexpERIgG %>% 
    mutate(phredAdjPval=-log10(adj.P.Val)) %>% 
    mutate(Signif=ifelse(adj.P.Val<=0.01, "<=0.01", ifelse(adj.P.Val<=0.05, "<=0.05", ">0.05"))) %>% 
    ggplot() +
        geom_point(aes(x=logFC, y=phredAdjPval, fill=Signif), colour="black", shape=21) +
        scale_fill_manual(values=c("red", "orange", "black"), name="Adjusted p-value", limits=c("<=0.01", "<=0.05", ">0.05")) +
        theme_minimal() +
        labs(x="log2(Fold Change)", y="-log10(Adjusted p-value)") +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(MAPlot, vp = define_region(row = 1, col = 1)) 
print(volPlot, vp = define_region(row = 1, col = 2))

```