### 2018-03-12 eRNA/plot_eRNA.Rmd
mRNA

---
title: "eRNA and mRNA"
author: "Andrew Holding"
date: "1/31/2018"
output:
  html_document: default
  pdf_document: default
---

```{r eRNAsetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Effect of GRHL2 overexpression on eRNA transcription

```{r eRNAlib, message=FALSE, warning=FALSE}
library("ggpubr")
```

```{r oeGRLH2, fig.cap="Effect of Overexpression of GRHL2 on eRNA at E2 responsive binding sites."}

eRNA<-read.csv("txt/eRNA_values.csv")

oe<-eRNA[eRNA$Experiment=="overexpression",]

oe_comparisons <- list( c("Control","GRHL2") )
oe_p <- ggboxplot(oe, x = "Treatment", y = "eRNA",
               add = "jitter", color="Treatment",
               facet.by="Gene", ylab="Relative eRNA levels")
oe_p <- oe_p + stat_compare_means(comparisons = oe_comparisons,
                       method = "wilcox.test" ,
                       paired=TRUE)
oe_p

```

## Effect of GRHL2 knockdown on eRNA transcription

```{r siGRHL2, fig.cap="Effect of knockdown of GRHL2 on eRNA at E2 responsive binding sites."}
si<-eRNA[eRNA$Experiment=="knockdown",]

si_comparisons <- list( c("siCtrl","siGRHL2") )
p <- ggboxplot(si, x = "Treatment", y = "eRNA",
               add = "jitter", color="Treatment",
               facet.by="Gene", ylab="Relative eRNA levels")
p + stat_compare_means(comparisons = si_comparisons,
                       method = "wilcox.test" ,
                       paired = TRUE)

```

###siRNA combined test
```{r wilcoxsiRNA}

wilcox.test(si$eRNA[si$Treatment=='siCtrl'],
            si$eRNA[si$Treatment=='siGRHL2'],
            paired=TRUE, alternative="less")

```

## Effect of GRHL2 overexpression on mRNA transcription

```{r mRNA}
qpcr<-read.csv("txt/QPCR GREB1 XBP1 TFF1.csv", header=TRUE)
df<-data.frame(qpcr)

p <- ggboxplot(df, x = "Sample", y = "Expression", ylab="Relative Expression",
               xlab="Condition",
               color = "Sample", palette =c("#00AFBB", "#E7B800", "#FC4E07"),
               add = "jitter", shape = "Sample", facet.by="Target",
               outlier=FALSE)

my_comparisons <- list( c("Ctrl", "siGRHL2"), c("Ctrl","GRHL2-DDK"))
p + 
    stat_compare_means(comparisons = my_comparisons,method="t.test", paired=TRUE)  +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```