### 2018-11-26 S6_eRNA/plot_eRNA.Rmd

---
title: "eRNA and mRNA"
author: "Andrew Holding"
date: "1/31/2018"
output:
  pdf_document: default
---

```{r eRNAsetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Effect of GRHL2 overexpression on eRNA transcription

```{r eRNAlib, message=FALSE, warning=FALSE}
library("ggpubr")
```

```{r oeGRLH2, fig.cap="Effect of Overexpression of GRHL2 on eRNA at E2 responsive binding sites."}

eRNA<-read.csv("txt/eRNA_values.csv")

oe<-eRNA[eRNA$Experiment=="overexpression",]

oe_comparisons <- list( c("Control","GRHL2") )
oe_p <- ggboxplot(oe, x = "Treatment", y = "eRNA",
               add = "jitter", color="Treatment",
               facet.by="Gene", ylab="Relative eRNA levels")
oe_p <- oe_p + stat_compare_means(comparisons = oe_comparisons,
                       method = "wilcox.test" ,
                       paired=TRUE)
oe_p

```

## Effect of GRHL2 knockdown on eRNA transcription

```{r siGRHL2, fig.cap="Effect of knockdown of GRHL2 on eRNA at E2 responsive binding sites."}
si<-eRNA[eRNA$Experiment=="knockdown",]

si_comparisons <- list( c("siCtrl","siGRHL2") )
p <- ggboxplot(si, x = "Treatment", y = "eRNA",
               add = "jitter", color="Treatment",
               facet.by="Gene", ylab="Relative eRNA levels")
p + stat_compare_means(comparisons = si_comparisons,
                       method = "wilcox.test" ,
                       paired = TRUE)

```

###siRNA combined test
```{r wilcoxsiRNA}

wilcox.test(si$eRNA[si$Treatment=='siCtrl'],
            si$eRNA[si$Treatment=='siGRHL2'],
            paired=TRUE, alternative="less")

```

## Effect of GRHL2  and GRLH2 delta 425-437 overexpression on eRNA transcription

```{r delta, fig.cap="Overexpression of GRHL2 delta 425-437."}
eRNA<-read.csv("txt/oe.csv")

eRNA<-eRNA[1:135,]

p <- ggboxplot(eRNA, x = "Condition", y = "Fold",
                   add = "jitter", color="Condition",
                   facet.by=c("Gene","CellTye"),ylab="Relative eRNA levels to empty vector")
comp.ev<-list(c("delta","OE"),c("delta","EV"),c("EV","OE"))
p2<-p + stat_compare_means(comparisons=comp.ev,
                           method = "t.test",paired=TRUE,alternative="greater",label.y=c(1.25,1.35,1.45))
p2

```